Sampling Strategies in Decisions from Experience

### 2022-05-20 manuscript.Rmd
Final Version: Section 1

The current paper commits to the core assumptions of the sequential sampling framework and uses its flexibility regarding the exact mechanisms of sampling, integration, and evaluation to incorporate the idea of sampling and decision strategies, including their interplay.
The two models presented below, one assuming an integration and evaluation mechanism resembling the round-wise strategy, the other assuming mechanisms resembling the summary strategy, offer a formal route to explore the potential interplay between sampling and integration/decision strategies, and their effects on the choice behavior.
The sequential sampling framework is chosen as a modeling outlet not only because it makes parsimonious assumptions and has assembled a good *"track record"* [cf. @meehlAppraisingAmendingTheories1990] by explaining and predicting actual choice behavior, but primarily because it makes explicit a core part of the computational problem the mind is trying to solve: 
in order to make systematic choices, the mind must assess differences in latent properties.
And in accumulating evidence for and against propsects, what should the mind accumulate evidence for if not such differences?
round-wise and summary decisions are both indicative for distinct properties the expected value and the probability to produce a higher outcome most of the time.
The remainder of the paper discusses and demonstrates in a simulation, how sampling and decision strategies as well as thresholds can act together to alter the degree and the accuracy with which these differences are assessed.

<!--Yet, there are several distinct properties that can be used to assess differences between the prospects, among them the EV, for which the summary decision strategy is indicative, and the probability to produce a higher outcome most of the time, for which the round-wise decision strategy is indicative [@hillsInformationSearchDecisions2010; see also @wulffHowShortLongrun2015].
Respectively, in the following, both decision strategies are implemented into a common sampling and accumulation model to demonstrate how the mind may rely on different sampling strategies to alter the degree to which evidence is accumulated for differences in either one of these properties.
More specifically, the proposed model accumulates the outcomes obtained from a sequence of comparisons between sample means, with the size of the samples underlying each mean comparison tending to be large (small) for low (high) switching probabilities.
Accordingly, whereas the decision variables of the model capture how many comparisons a prospect has won, the unfolding accumulation process more closely resembles the summary strategy for low switching probabilities and the round-wise strategy for high switching probabilities.

Now, the current paper presents a sampling and accumulation model that provides an algorithmic link between sampling and decision strategies, and uses the model to discuss how the mind may rely on sampling strategies to alter which differences in the latent properties of prospects evidence is accumulated for.
More specifically, the proposed model assumes an accumulation process over a sequence of mean comparisons, with switching probabilities---i.e., the probability of sampling
outcomes from different prospects in direct succession---determining the size of the samples underlying each mean comparison and thresholds determining how many comparisons a prospect must win in order to be chosen.
The model is such that when switching probabilities are high, the accumulation process unfolds over mean comparisons based on small samples, thus resembling the round-wise strategy and accumulating evidence for differences in the probability to produce a higher outcome most of the time.
Notably, in this case, small-probability outcomes are predicted to contribute to the majority of mean comparisons less than would be objectively warranted, which can translate to an as-if underweighting of rare outcomes, irrespective of what the thresholds require the number of won comparisons to be.
In turn, for low switching probabilities, the accumulation process unfolds over mean comparisons based on large samples, thus resembling the summary strategy---for low thresholds, respectively---and accumulating evidence for differences in the EV.-->

>>>>

The current paper commits to the core assumptions of the sequential sampling framework and uses its flexibility regarding the exact mechanisms of sampling, integration, and evaluation to incorporate the idea of sampling and decision strategies, including their interplay.
The two models presented below, one assuming an integration and evaluation mechanism resembling the round-wise strategy, the other assuming mechanisms resembling the summary strategy, offer a formal route to explore the potential interplay between sampling and integration/decision strategies, and their effects on the choice behavior.
The sequential sampling framework is chosen as a modeling outlet not only because it makes parsimonious assumptions about the decision-making process and accounts for the behavior in various decision-making tasks [@ratcliffComparisonSequentialSampling2004], but primarily because it makes---taking a Marrian perspective [cf. @marrVisionComputationalInvestigation1982; @marrUnderstandingComputationUnderstanding1977; see also, e.g., @griffithsProbabilisticModelsCognition2010]---explicit a core part of the computational problem the mind is trying to solve in DFE, and the principle logic that governs the solution:
In order to make systematic choices between prospects, the mind must assess true differences in their *latent* properties, and it can only do so by searching for evidence for such differences in information samples.
Sequential sampling models of DFE make this clear by assuming an evidence accumulation process over sampled outcomes that is more or less directed towards particular latent properties.
Round-wise and summary decisions are both indicative for distinct latent properties, i.e., the expected value and the probability to produce a higher outcome most of the time, respectively.
The remainder of the paper discusses and demonstrates in a simulation, how integration/decision strategies can act together with sampling strategies and evidence thresholds to alter the degree and the accuracy with which differences about either one of these properties are assessed.
